\section{Motivation and the gap}

We want to predict which residues of an antigen are B-cell epitope residues, from the antigen
structure alone, in cases where the antibody is unknown. Doing this rigorously requires
residue-level labels for ``binding importance.'' We take these from \textbf{alanine scanning}:
for residue $i$, $\Delta\Delta G_\mathrm{bind}^{(i)} = \Delta G_\mathrm{bind}^{\mathrm{mut},i} -
\Delta G_\mathrm{bind}^{\mathrm{WT}}$, with large positive values flagging residues whose
wild-type identity stabilizes the interface. We treat this as a residue-level epitope score,
$S_i^\mathrm{epitope} \approx \Delta\Delta G_\mathrm{bind}^{(i)}$. (Full derivation and the
alchemical / StaB-ddG label tiers: \texttt{../../tcr/docs/b\_cell\_eptiope\_prediction\_may23\_update\_final.pdf}.)

\paragraph{The gap.} Most B-cell epitope predictors --- and the benchmarks that train them ---
label residues by \emph{contact distance} to a bound antibody. As a coarse first pass this is
defensible: proximity to the paratope is a reasonable way to \emph{nominate} candidate interface
residues. The limitation is treating every contact as a positive label of equal weight, which
conflates ``close'' with ``important'': many interface residues contribute little binding free
energy, while a few hotspots carry most of the affinity. The bias is baked into the data itself.
Of the $4{,}164$ conformational (discontinuous) B-cell epitopes in IEDB that carry a deposited
structure, essentially all ($4{,}161$; $>99\%$) are defined by crystallographic or cryo-EM
\emph{contact} rather than by any binding measurement;\footnote{IEDB next-generation query API
(\texttt{query-api.iedb.org}), \texttt{bcell\_search}, structure type \emph{Discontinuous peptide}
with a non-null \texttt{pdb\_id}, queried 2026-06. IEDB attaches a structure only when the epitope
was itself defined from one, so structure-linked records are overwhelmingly contact-derived; the
mutation/binding-derived epitopes carry no \texttt{pdb\_id}.} models trained on these labels
therefore inherit proximity as their working definition of ``epitope.'' Alanine-scan
$\Delta\Delta G$ supplies the missing energetic refinement --- \emph{which} of the contacting
residues actually pay for binding. Even a binary critical/non-critical call from a saturating
alanine scan is energetically grounded, since ``critical'' means mutation measurably weakens
binding; proximity and energetics are thus complementary, the former a candidate filter and the
latter the importance label.

Separately, since Westhof et al.~\cite{westhof1984} it has been known that antigenicity correlates
with \emph{segmental mobility} (flexibility / B-factor) --- an antigen-intrinsic, antibody-agnostic
property --- and energy-decomposition analyses of plain MD trajectories can already localize
antibody epitopes on an antigen without knowledge of the antibody~\cite{serapian2020}. The
opportunity is to test, with modern MD, whether antigen-only \emph{dynamical} and \emph{solvation}
features (RMSF, collective-mode participation, hydration thermodynamics) predict energetic binding
importance --- and whether they beat static exposure. This is where the author's methods background
(FRESEAN collective-mode analysis; 3D-2PT-style hydration thermodynamics; non-equilibrium energy
flow) is the differentiator (\texttt{../../\_literature/my\_manuscripts/}).
