\section{Methods --- apo-HEL run \#1}

Frozen, signed-off stack in \texttt{../docs/METHODS.md} (2026-06-18). Purpose of run \#1:
\emph{validate the prep$\to$equil$\to$production pipeline on Gemini and extract a first RMSF /
flexibility feature} (the Westhof re-test). House default stack signed off as the starting point.

\paragraph{Stack.} AMBER99SB-ILDN / TIP3P; cubic box, 1.2\,nm clearance; neutralize $+$
0.15\,M NaCl; leapfrog, 2\,fs, LINCS on h-bonds; PME ($r_\mathrm{coul}=r_\mathrm{vdw}=1.0$\,nm,
potential-shift, DispCorr); V-rescale at 300\,K (Protein / Non-Protein); EM $\to$ 100\,ps NVT $\to$
100\,ps NPT (both position-restrained) $\to$ \textbf{100\,ns} production (single trajectory, for
RMSF statistics). Random seeds (\texttt{gen\_seed} $=$ \texttt{ld\_seed} $= -1$): bitwise
reproducibility is not the target --- the workflow reproduces from git-tracked inputs and saved
derived outputs, and random seeds give genuinely independent replicas. Frozen per-run \texttt{.mdp}
files: \texttt{md/1AKI/apo/configs/} (run dirs are organized \texttt{md/<PDB>/<state>/}).

\paragraph{Deliberate deviations from the house scripts.}
(i) \textbf{C-rescale} barostat for NPT equilibration (Parrinello-Rahman only in production), since
PR can ring when started far from equilibrium; (ii) \textbf{cubic} box (the old house scripts
disagreed cubic/dodecahedron) --- chosen for interpretability and uniform across all systems, at a
modest extra-solvent cost vs.\ dodecahedron.

\paragraph{HEL-specific checkpoints (not glossed over).} pdb2gmx must form HEL's \textbf{4 disulfides}
(6--127, 30--115, 64--80, 76--94) --- verify the log reports 4 SS bonds. His15 protonation: record the
assigned state. Crystal waters (78 HOH) removed; fresh TIP3P added. Read every grompp warning rather
than letting \texttt{-maxwarn} hide it.

\paragraph{Caveats / flags.} TIP3P over-mobilizes water and 2\,fs$+$LINCS suppress H dynamics --- fine
for relative RMSF and pipeline validation, \emph{not} for the dynamics-grade features. 100\,ns is a single
trajectory for statistics; replicas are deferred (random seeds make adding independent ones trivial).

\paragraph{Deferred: the distinctive run.} FRESEAN mode participation and 3D-2PT hydration
thermodynamics require a \emph{separate} protocol --- high-frequency velocity output, constraint
removal / sub-fs dt, TIP4P/2005 --- with settings pulled from the FRESEAN and energy-transfer
manuscripts. This is the run that powers the mode-based and hydration feature classes.

\section{Structure completeness and the repair strategy}\label{sec:repair}

Several crystal structures are incomplete --- disordered termini, internal loops, or whole
segments that the experiment never resolved (PDB \texttt{REMARK 465}). Apo-MD demands a chemically
complete chain, so each gap forces a choice. We make that choice explicitly per structure rather
than feeding everything through one tool, because \textbf{the right repair scales with gap length},
and the wrong tool on a long gap produces a confident-looking but meaningless model.

\paragraph{What a ``fixer'' can and cannot supply.} The instinct to ``just grab the FASTA for the
missing residues'' conflates two different problems. The deposited sequence (\texttt{SEQRES}, the
same residues a fixer recovers automatically by diffing \texttt{SEQRES} against the resolved
\texttt{ATOM} records) supplies the \emph{identity} of the missing residues --- but identity is the
easy half. The hard half is \emph{geometry}: where the missing backbone atoms sit in three
dimensions, about which the sequence says nothing. \texttt{pdbfixer} (OpenMM) builds missing
heavy atoms onto resolved residues from templates --- valid and routine --- and threads missing
residues in with default internal coordinates followed by a short minimization of only the new
atoms. That minimization removes clashes; it performs \emph{no} conformational sampling. The result
is therefore reliable for a short surface gap (apo equilibration relaxes a two- or three-residue
loop anyway) and arbitrary for anything longer, where real loop modelling (\textsc{Modeller},
Rosetta) or a full predicted model (AlphaFold) is required. Disordered \emph{termini} are not
modelled at all: they bear no labels and would only add unrestrained flailing, so we cap them.

\paragraph{Method used: AlphaFold-loop grafting with restrained closure.} Rather than let a
fixer build loops blind, we graft the missing internal residues from an AlphaFold model and use
the crystal for everything resolved. For each antigen we (i) sequence-align the crystal chain to
its AlphaFold model to fix the numbering offset, (ii) measure the C$\alpha$-RMSD between model and
crystal over all matched residues --- the ``is the prediction trustworthy here'' test --- and
(iii) for each internal gap, locally superpose the model on the flanking resolved residues and
splice in its loop at crystal numbering. \texttt{pdbfixer} then completes any partial side chains,
and a short OpenMM minimization with all crystal atoms position-restrained (only the loop free,
plus an explicit harmonic across the spliced junction) closes the peptide bonds. Disordered termini
are capped, not modelled. Scripts: \texttt{analysis/repair\_structure.py},
\texttt{finalize\_structure.py}, \texttt{close\_loops.py}; host-side env
\texttt{env/repair-environment.yml} (Python~3.11, OpenMM~8.5, \texttt{pdbfixer}, run locally, not on
Gemini).

\begin{center}\small
\begin{tabular}{@{}l p{0.34\linewidth} l@{}}
\toprule
Antigen (PDB, chain) & Unresolved segments (\texttt{REMARK 465}) & Repair outcome \\
\midrule
hGH (\texttt{1HGU} A) & F1; \textbf{K38--E39} (internal); F191 & AF-graft (AFDB \texttt{P01241}); \textbf{done} \\
Tissue factor (\texttt{1AHW} C) & S1--T3; \textbf{V83--G90} (8-res); Q212--E219 & AF-graft (AFDB \texttt{P13726}); \textbf{done} \\
HCMV gB (\texttt{5C6T} A) & 63--87 (25); 115--120; 219--220; 237--239; \textbf{437--468 (32)} & not in AFDB $\to$ AF3 prediction (pending) \\
\bottomrule
\end{tabular}
\end{center}

\noindent For tissue factor the AlphaFold model agrees closely with the crystal (global
C$\alpha$-RMSD $1.12$\,\AA\ over 200 residues; the V83--G90 loop fits its flanks at $1.29$\,\AA),
so the 8-residue graft is well supported. For hGH only the 2-residue K38--E39 turn is grafted
(local fit $1.81$\,\AA); the larger global RMSD ($3.36$\,\AA, hinge flexibility of the four-helix
bundle) is immaterial for so short a splice. Both repaired chains validate as continuous, with the
expected disulfides (hGH 53--165, 182--189; TF 49--57, 186--209) and no steric clashes. HCMV gB is
a viral protein \emph{absent from AlphaFold~DB}, and its largest gap (32 residues) is true
\emph{de novo} prediction, not loop repair; it requires a fresh model. We generate one with
\textbf{AlphaFold3} on Gemini (the lab's \texttt{alphafold\_3.0.1} Singularity container and
installed weights/databases; \texttt{md/5C6T/apo/af3.sbatch}, mirroring the episcaf pipeline), then
graft its loops the same way. The AF3 model's agreement with the crystal over the resolved residues
is the same trust check applied above; if it disagrees we simulate the AF3 ectodomain directly.

\paragraph{Consequences for the run queue.} Eight of the nine non-pristine/pristine antigens are
now ready to prep on Gemini (\texttt{git pull} then \texttt{sbatch}): pristine chains HEL
(\texttt{1AKI}), HPr (\texttt{2JEL}), VEGF (\texttt{1BJ1}), IFN-$\gamma$R (\texttt{1JRH}),
Dengue~E (\texttt{1OKE}), HA (\texttt{1HGG}); plus the repaired hGH (\texttt{1HGU}) and tissue
factor (\texttt{1AHW}). Outstanding: HCMV gB (\texttt{5C6T}, awaiting an AF3 model) and the
previously deferred MT-SP1 (\S\ref{sec:simulated-antigens}, two internal residues, same AF-graft
recipe) and Bont/A1.
