\section{Run log}

Exact-command reproducibility ledger. Append a row per stage; pin the GROMACS version actually
used on Gemini, seeds, and Slurm job id.

\begin{center}
\begin{tabular}{lllll}
\toprule
Date & Stage & Command & GROMACS ver. & Job / output \\
\midrule
2026-06-18 & structure & \texttt{curl .../1AKI.pdb} & --- & \texttt{structures/1AKI.pdb} \\
\textit{pending} & prep & \texttt{sbatch md/1AKI/apo/prep.sh} & 2023.2-dev & out/npt.gro/cpt \\
\textit{pending} & production & \texttt{sbatch md/1AKI/apo/md.sh} & 2023.2-dev & out/md.xtc \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Status (2026-06-18).} Project scaffolded; benchmark audited; first system (HEL/1AKI)
selected and label mapping verified 1:1; apo-HEL run \#1 staged and signed off. Gemini
environment confirmed: module \texttt{Gromacs} (capital G), GROMACS \texttt{2023.2-dev}
(avx2\_256), partition \texttt{gpu-a100} (4-day limit, A100 fastest); repo cloned to
\texttt{/scratch/bneff/bcell\_epitope}. Production walltime set to 48\,h ceiling.
\textbf{Next:} \texttt{git pull} on Gemini, then \texttt{sbatch prep.sh} $\to$ \texttt{sbatch
md.sh}; record Slurm job ids and ns/day. Production length 100\,ns (single trajectory). Open:
replicas (deferred --- random seeds make them trivial); scheduling the FRESEAN/3D-2PT run.

\paragraph{Run-script fixes (bringing up run \#1 on Gemini).} Three issues surfaced and were
fixed while getting prep to run; recorded for reproducibility:
\begin{itemize}
  \item \textbf{Module name.} Gemini's module is \texttt{Gromacs} (capital G); lowercase
        \texttt{gromacs} errors on a missing \texttt{shared} dependency. Scripts updated.
  \item \textbf{Config paths under Slurm.} Slurm copies the batch script to a private spool
        dir, so \texttt{\$(dirname \$0)/configs} resolved into the spool dir and \texttt{ions.mdp}
        was not found. Scripts now anchor paths to \texttt{\$SLURM\_SUBMIT\_DIR};
        \textbf{always submit from the repo root}.
  \item \textbf{GPU bonded offload on EM.} \texttt{-bonded gpu} requires a dynamical integrator;
        EM uses \texttt{steep}, so the EM \texttt{mdrun} now uses \texttt{-nb gpu} only. NVT/NPT/
        production use \texttt{md} and keep \texttt{-nb/-pme/-bonded gpu} (audited).
\end{itemize}
Also removed the obsolete \texttt{ns\_type} entry from all mdp files (ignored under the Verlet
scheme). \textbf{Caveat:} Gemini scratch is not durable --- move logs/outputs off scratch after
each run (an earlier prep's logs were lost to a purge). Sanity check per run: \texttt{pdb2gmx.log}
must report \textbf{4 disulfides} for HEL.

\paragraph{Expansion to multiple antigens (2026-06-21).} Switched the box from dodecahedron to
\textbf{cubic} (interpretability; uniform across systems) and re-ran HEL for consistency. Staged
three further apo-antigen systems from the benchmark, hybrid-sourced (strip antigen chain(s) from
the complex; numbering kept aligned to the labels). Provenance:
\begin{itemize}
  \item \textbf{HPr} --- \texttt{2JEL} chain P (res 1--85, no Cys); \texttt{md/2JEL/apo}.
  \item \textbf{VEGF} --- \texttt{1BJ1} chains V$+$W (res 14--107, disulfide-linked homodimer;
        \texttt{pdb2gmx -merge all} for the inter-chain SS); \texttt{md/1BJ1/apo}.
  \item \textbf{Bont/A1 Hc} --- \texttt{2NYY} chain A res 872--1295 (C-terminal receptor-binding
        domain only; catalytic-domain Zn/Ca excluded; res 872 is a domain cut $\to$ charged
        N-terminus); \texttt{md/2NYY/apo}.
\end{itemize}
All four (HEL$+$these) submitted overnight as prep$\to$production chains
(\texttt{sbatch --dependency=afterok}). \textbf{Deferred:} MT-SP1 (2 missing internal residues),
Bont/A2 (homology models), hGH (provisional labels). Per-system disulfide
expectations recorded in each \texttt{prep.sh}.

\paragraph{Structure repair (2026-06-29).} Audited \texttt{REMARK 465} for the incomplete staged
structures and repaired two of three by AlphaFold-loop grafting with restrained closure (method +
RMSDs in \S\ref{sec:repair}; scripts \texttt{analysis/repair\_structure.py},
\texttt{finalize\_structure.py}, \texttt{close\_loops.py}; host env
\texttt{env/repair-environment.yml}). \textbf{hGH} (\texttt{1HGU}, K38--E39 grafted from AFDB
\texttt{P01241}) and \textbf{tissue factor} (\texttt{1AHW}, loop V83--G90 from AFDB
\texttt{P13726}; AF--crystal global C$\alpha$-RMSD 1.12\,\AA) are now frozen as
\texttt{structures/\{hGH\_1HGU,TissueFactor\_1AHW\}\_fixed.pdb} --- continuous, correct disulfides,
clash-free --- so both are ready to prep. \textbf{HCMV gB} (\texttt{5C6T}) is \emph{absent from
AlphaFold~DB} (viral) and its 32-residue gap is de-novo prediction; deferred pending an AF3
(AlphaFold Server) model. Net: eight antigens ready to \texttt{sbatch} on Gemini.

\paragraph{Dynamics-feature push (2026-06-30 / 07-01) --- what happened, including the wrong turns.}
Two days of apo-MD feature extraction and scoring against the \emph{energy} label (alanine-scan
$\ddG$), on 6 Tier-1 antigens (\texttt{1AKI 2JEL 1BJ1 1JRH 1AHW 1HGU}; $\sim$9 antibody-epitopes,
$\sim$97 scanned residues). Features computed on Gemini: RMSF, trajectory SASA and SASA fluctuation,
slow-mode (positional-PCA) participation, DCCM coupling, $\phi/\psi/\chi_1$ dihedral flexibility,
per-residue H-bond counts (intra-protein and to water), and gRINN intra-protein interaction energy.
Evaluation: \emph{surface-restricted directional AUROC} --- restrict to surface residues (trajSASA
above the per-antigen median), then $P(\text{feature}_{\text{hotspot}} > \text{feature}_{\text{rest}})$,
reporting the sign; leave-one-antigen-out for generalisation. Under it, raw exposure (SASA) collapses
to $\approx 0.51$ (chance), confirming the baseline the project is built to beat.

\emph{Four bugs surfaced, each of which moved a headline number:} (i) the alanine-scan $\ddG$ table is
in antibody-\emph{complex} PDB numbering while the apo-MD structures are numbered differently --- every
feature$\leftrightarrow\ddG$ join on raw residue number mislabelled 4 of 6 systems (only $39/101$
labelled residues matched by wild-type identity; $101/101$ after per-system offsets
\texttt{1AHW}$-3$, \texttt{1BJ1}$-13$, \texttt{1HGU}$-1$, \texttt{1JRH}$-10$; persisted in
\texttt{benchmark/labels/numbering\_offsets.json} + \texttt{ddg\_by\_md\_residue.csv}, WT-verified).
(ii) gRINN's output is $0$-indexed while every other file is $1$-indexed --- the energy join was off by
one residue; correcting it (\texttt{grinn.resid $+=1$}) \textbf{collapsed interaction-energy detection
$0.67\to0.57$}: the apparent energy signal, and the ``internal-energy / protrusion / under-satisfied
side-chain'' mechanism built on it, were an alignment artefact --- \textbf{that narrative is dead}.
(iii) gRINN for VEGF ran on a single protomer of the obligate homodimer (\texttt{1BJ1}, 94 vs 188
residues) --- energy scale and PC1 broken; \texttt{1BJ1} excluded. (iv) reproducibility gaps: the
reassessment computation lived only in an ephemeral session (committed script was a stub), the two
figures were not regenerable, and the committed trajSASA carried \texttt{NaN} for \texttt{UNK}
residues. A runnable evaluation script now replaces the stub
(\texttt{analysis/reassess\_corrected\_labels.py}): it applies the offsets with a WT-identity assertion,
fixes the gRINN index in one place, and locks the numbers.

\emph{Corrected, reconciled headline} (5 systems, \texttt{1BJ1} dropped; surface, per-antigen AUROC):
H-bonds to water $0.66$ (best single feature, survives every correction), H-bonds intra $0.60$,
trajectory SASA $0.57$, interaction energy (vdW) $0.57$, DCCM $0.56$, backbone flexibility $0.52$,
PC1 participation $0.50$, SASA fluctuation $0.40$. The HB-water \emph{gradient} among labelled residues
is $\rho=-0.33$ (an earlier $-0.50$ was wrongly surface-restricted, dropping \texttt{2JEL}, which runs
the other way, $\rho=+0.27$); it is not unanimous. \textbf{Honest state:} no single antigen-only feature
is a strong predictor of the energy label at $N\approx6$; every exposure/flexibility feature is weak
(as the thesis predicts), and interaction energy is now dead. The real bottleneck is $N$ --- no
transferable linear combination beats the best single feature (LOAO logistic goes below chance) ---
which motivates scaling the label set. \emph{Still open:} reconcile the HB-water detection number
between the runnable script ($\approx0.52$ on the shared surface set) and the Science-kernel $0.656$
(a residual AUROC/negative-set difference); regenerate the feature figures once locked; the
FRESEAN low-frequency-mode run (proper replacement for the PCA placeholder) is in progress.
\emph{Note for a later revision:} the draft energetics narrative rests on the now-retracted energy
result and needs rewriting once FRESEAN and the label scale-up land --- flagged, not yet edited.

\medskip
\emph{Session \#5 (2026-07-07), collective modes + strict evaluation (\S\ref{sec:modes}).} FRESEAN
low-frequency modes landed for \texttt{1AKI} (the proper replacement for the PCA placeholder). The
promising single-system involvement signal ($\mathrm{AUROC}=0.75$) is a buried/exposed confound: surface-
restricted it falls to $0.64$ (n.s.), isolation to chance and flipped (integration-vs-SASA $-0.68$). The
``co-moving decoupled patch'' is generic --- all 12 dynamical communities show it, epitope patches no more
than non-epitope ($0.55$ vs $0.56$). Colombo's MLCE tested on its own gRINN interaction-energy matrix:
qualitative overlap on 2/6 systems, strict surface AUROC $0.49$, and $-0.84$ redundant with the interaction
energy already in the set (zero marginal LOAO gain). Exhaustive 2047-subset search maxes at $0.61$ (inside
the shuffled null); MLP memorises ($0.58\pm0.03$ test). \textbf{The decisive test} --- train on 6 Tier-1
antigens, predict unseen Dengue E ($n/N$ shotgun label): every physics-feature set at chance (0.43--0.52),
only exposure transfers ($0.58$, $\rho=+0.15$). \textbf{The physics-feature program is now closed at $N=6$:}
flexibility, energetics, dynamical and energetic coupling do not beat exposure under leakage-controlled,
cross-antigen, per-residue evaluation. This is the checkpoint that motivates the scale-up --- SKEMPI label
expansion (211 true antigen-side labels / 19 complexes) and the \texttt{1OKE} dense-label FRESEAN run
(chunk 1 validated, 2--20 pending) --- rather than a fancier model. Per-antibody lysozyme decomposition
(from SKEMPI, zero new MD) is the one clean positive: two antibodies map to two dynamically distinct patches.
